\documentclass[a4paper]{article}

% Expanded on 2022-05-05 at 14:21:39.

\usepackage{../../style}

\title{AICC2}
\author{Joachim Favre}
\date{Jeudi 05 mai 2022}

\begin{document}
\maketitle

\lecture{20}{2022-05-05}{Minimum distance}{
}

\parag{Definition: Minimum distance decoding}{
    For any code $\mathcal{C}$ and for any channel yielding $y$, the Minimum Distance (MD)-Decoder outputs:
    \[\hat{c} \in \arg \min_{x \in \mathcal{C}} d_H\left(y, x\right)\]

    The arg maps back to the codeword. We are having a set since there might be multiple minimums.

    In other words, the decoder makes a table, measuring the distance between what it gets and all the possible codewords. Then, it just picks the one with the smallest Hamming distance.
}

\parag{Definition: Minimum distance of a code}{
    Let's say we have a code $\mathcal{C}$. Then, the \important{minimum distance} of this code, is defined as:
    \[d_{min}\left(\mathcal{C}\right) \over{=}{def} \min_{\substack{x, y \in \mathcal{C} \\ x \neq y}} d_H\left(x, y\right)\]
}

\parag{Theorem: Erasure channel}{
    Let's consider an erasure channel.

    If the erasure weight is such that $p < d_{min}\left(\mathcal{C}\right)$, then the MD-decoder is guaranteed to find the correct codeword.

    \subparag{Remark}{
        The greatest $p$, $p_{max}$, such that for any erasure weight $p \leq p_{max}$, we can be certain that the correct codeword can be found, is given by:
        \[p_{max} = d_{min} - 1\]
    }

}

\parag{Theorem: Error channel}{
    Let's consider an error channel.

    If the error weight is such that $p < \frac{d_{min}\left(\mathcal{C}\right)}{2}$, then the MD-decoder is guaranteed to find the correct codeword.


    \subparag{Remark 1}{
        The greatest $p$, $p_{max}$, such that for any error weight $p \leq p_{max}$, we can be certain that the correct codeword can be found, is given by:
        \[p_{max} = \left\lfloor \frac{d_{min} - 1}{2} \right\rfloor\]
    }

    \subparag{Remark 2}{
        The greatest $p$, $p_{max}$, such that for any error weight $p \leq p_{max}$, we can be certain that an error \textit{can be detected} is given by:
        \[p_{max} = d_{min} - 1\]

        Indeed, if $p \leq d_{min} - 1$, then we cannot mistake what we get with another codeword.
    }

}

\parag{Theorem: Singleton's bound}{
    For any code $\mathcal{C}$, regardless of the alphabet $\mathcal{A}$, we have:
    \[d_{min}\left(\mathcal{C}\right) \leq n -k + 1\]
    where $n$ is the block length and $k = \log_{\left|\mathcal{A}\right|}\left(\left|\mathcal{C}\right|\right)$.

}

\parag{Definition: MDS code}{
    A code $\mathcal{C}$ that attains Singleton's Bound with equality is called \important{Maximum Distance Separable (MDS) code}.
}

\end{document}
